### 2021-06-15 JSS/using_drtmle_jss.Rmd
final touch-ups

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
    orcdi: 0000-0002-1019-8343
  - name: Nima S. Hejazi
    affiliation: University of California, Berkeley
    email: \email{nhejazi@berkeley.edu}
    orcdi: 0000-0002-7127-2789
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \pkg{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect. The package additionally implements a targeted inverse probability of
  treatment weighted estimator that yields valid inference even when the
  probability of treatment is estimated via adaptive methods.
header-includes:
  - \usepackage{mathtools}
output: rticles::jss_article
bibliography: refs
---

@@

The two proposed procedures that yield doubly-robust inference are quite
involved, notably requiring iterative estimation of additional, complex
nuisance parameters. The `drtmle` package was motivated by the need for a
user-friendly tool to implement these methods. The package implements the TML
estimators of population-level effects proposed in \citet{vdl:2014:ijb} and
\citet{benkeser:2017:biometrika}, as well as several extensions including
cross-validated targeted minimum loss-based estimation (CV--TMLE). Additionally,
`drtmle` implements an IPTW estimator that overcomes shortcomings of existing
IPTW implementations with respect to statistical inference. In particular, the
estimator allows an adaptive estimator of the PS to be used in construction of
the IPTW estimator, while yielding valid statistical inference about
population-level effects. Both the TML and IPTW estimators implemented in the
`drtmle` package are capable of estimating population-level effects for
discrete-valued treatments; moreover, the package provides convenient utilities
for constructing confidence intervals and hypothesis tests about contrasts of
the mean outcome under different levels of treatment. This document explains
some of the key concepts underlying doubly-robust inference and illustrates
usage of the `drtmle` package.

@@

The main function of the package is the eponymous `drtmle` function. This
function estimates the treatment-specific marginal mean for user-specified
levels of a discrete-valued treatment and computes a doubly-robust covariance
matrix for these estimates. The function implements either the
estimators of \citet{vdl:2014:ijb} (if `reduction = "bivariate"`) or the
improved estimators of \citet{benkeser:2017:biometrika}
(`reduction = "univariate"`). The package includes utility functions for
combining these estimates into estimates of average treatment effects, as well
as other contrasts, as discussed below.

The main data arguments for `drtmle` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
regression parameters may be estimated. Once the user has specified the data
arguments and how to estimate each regression parameter, the function proceeds
as follows:
  1. estimate the OR via the internal `estimateQ` function;
  2. estimate the PS via the internal `estimateG` function;
  3. estimate reduced-dimension regression via the internal `estimateQrn` and
     `estimategrn` functions;
  4. iteratively modify initial estimates of the OR and PS via the `fluctuateQ`
      and `fluctuateG` functions until equations (5)-(7) are approximately
      solved;
  5. compute the TML estimates and covariance estimates based on the final OR.
We refer interested readers to \citet{benkeser:2017:biometrika} for a
theoretical derivation of how the iterative modification of initial OR and PS is
performed.

In addition to returning doubly-robust parameter and covariance estimates, the
`drtmle` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly-robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we look at various
options for making calls to `drtmle`.

## Estimating regressions
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated, then the estimators
computed by `drtmle` achieve the nonparametric efficiency bound, resulting in
narrower confidence intervals and more powerful hypothesis tests. In order for
the estimates of the OR and PS to have the greatest chance of consistency, the
`drtmle` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning
[@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
technique. We demonstrate each of these approaches in turn on a simple simulated
data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
2)` and $\psi_0(1)=$ `r round(EY1, 2)`.

@@

The code below illustrates how generalized linear models can be used to
estimate the regression parameters. In this call to `drtmle`, we specify
`family = binomial()` so that logistic regression is used to estimate the OR.
We illustrate non-stratified regression for both the univariate (default) and
bivariate reductions below.

@@

Super learning is implemented in the R package \pkg{SuperLearner}
[@SuperLearnerPackage] and `drtmle` provides the option to utilize super
learning for estimation of the OR, PS, and reduced-dimension regressions via
the `SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle` either
the generalized linear model options for regression modeling or the super
learner options should be invoked. If both are called, the function will
automatically opt to using the super learner option. The inputs to the super
learner options are passed to the `SL.library` argument of the `SuperLearner`
function of the `SuperLearner` package. We refer readers to the `SuperLearner`
documentation for information on how to properly specify a super learner
library. In the code below, we illustrate a call to `drtmle` with a relatively
simple library. For the OR and PS, the library includes a main-terms generalized
linear model (via the function `SL.glm` from `SuperLearner`) and an unadjusted
generalized linear model (`SL.mean`). For the reduced-dimension regressions we
use a super learner library with a main-terms generalized linear model
(`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
`earth` package [@earthPackage]).

@@

The `drtmle` package also allows users to pass in their own functions for
estimating regression parameters via the `SL_` options. These functions must be
formatted properly for compatibility with `drtmle`. The necessary formatting is
identical to the formatting required for writing a `SuperLearner` wrapper (see
`SuperLearner::write.SL.template()`). The `drtmle` package includes `SL.npreg`,
which is an example of how a user can specify a function for estimating
regression parameters. This functions fits a kernel regression estimator using
the `np` package [@npPackage]. Below we show the source code for this function.

@@

In the above code, we see that a user-specified function should have options
`Y, X, newX, family, obsWeights, ...`, along with other options specific to the
function. The option `Y` corresponds to the outcome of the regression and `X`
to the variables onto which the outcome is regressed. The option `newX` should
be of the same class and dimension as `X` and is meant to contain new values of
variables at which to evaluate the regression fit. The `family` and
`obsWeights` options are not strictly necessary (indeed, they are not used by
`SL.npreg`), though these may be useful to include if one wishes to utilize the
function as part of a super learner library. The `SL.npreg` function proceeds by
checking whether there is sufficient variation in `Y` to justify fitting the
kernel regression (as specified by option `rangeThresh`), and if so, fits a
kernel regression with cross-validated bandwidth selection via the `npregbw()`
function. The output is then formatted in a specific way so that `drtmle` is
able to retrieve the appropriate objects from the fitted regression. In
particular the output should be a list with named objects `pred` and `fit`; the
former including predictions made from the fitted regression on `newX`, the
latter including any information that may be needed to predict new values based
on the fitted object. A class is assigned to the `fit` object, so that an S3
`predict` method may later be used to predict new values based on the fitted
regression object. The user must also specify this `predict` method. Use
`drtmle:::predict.SL.npreg` to view the simple `predict` method for `SL.npreg`.

In the code below, we demonstrate a call to `drtmle` with `SL.npreg` used to
estimate each nuisance parameter. Here, we also illustrate the use of `stratify
= TRUE`, which fits a separate kernel regression of $Y$ onto $W$ in
observations with $A = 1$ and $A = 0$.

@@

Not surprisingly, averaging over multiple super learners can add considerable
computational expense to calls to `drtmle`. The package is currently
parallelized at the level of individual calls to `SuperLearner`. With
parallelization over enough cores, repeated super learning should not add too
much to the computational burden relative to a single call to `SuperLearner`.

@@

The `drtmle` package also allows users to pass in their own estimated OR and PS
via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the user,
`drtmle` will ignore the nuisance parameter estimation routines specified by
`SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As noted in the
package documentation, there is a specific format necessary for inputting `Qn`
and `gn`. For `Qn`, the entries in the list should correspond to the OR
evaluated at `A` and the observed values of `W`, with order determined by the
input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the OR
at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
example below illustrates this concept.

@@

This section contains some more advanced mathematical material and may be
skipped over by users interested only in implementation. For simplicity, we focus
this discussion on the standard TML/AIPTW estimators. The ideas generalize
immediately to the DR-TML estimator, but the formulas are more complex in this
latter case.

@@

For TML estimators, it is standard practice to construct $D_i$ using the
*targeted* version of the OR (denoted by $\bar{Q}_n^{\star}$ above). Recall that
such estimates are obtained by modifying an initial OR estimate so as to ensure
a particular (set of) equation(s) is (are) solved.  However, simulations have
shown that better standard error estimates for TMLE may be achieved by instead
using the *initial* OR estimate in the computation of standard errors, as
described in the preceding subsection. To this end, The `drtmle` function
includes the `targeted_se` option.

@@

Suppose we use $V$-fold cross-validation to obtaine $V$ training-sample-specific
estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the
estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the
validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of
$Y_i$. We define $g_{n,i}^{cv}$ similarly. Using these estimates, we can define
\begin{equation*}
 D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} +
 \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1)
\end{equation*}
and compute an estimate of $\sigma^2_0$ using
\begin{equation*}
  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} -
  \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ .
\end{equation*}

In `drtmle`, cross-validated standard error estimates can be obtained by setting
`se_cv = "full"` and `se_cvFolds` to a number greater than 1. For example,

@@

Obtaining fully cross-validated standard errors can be fairly time-intensive
when super learner is used to estimate the OR and PS. In `drtmle`, we give an
option to compute standard error estimates that are *partially cross-validated*
and do not require any additional fitting beyond a standard super learner. Here,
we describe briefly how these estimates are obtained and show calls to `drtmle`
for obtaining such estimates.

@@

Below, we demonstrate how to obtain such estimates using `drtmle`.

@@

The `drtmle` package contains methods that compute doubly-robust confidence
intervals. The `ci()` method computes confidence intervals about the estimated
marginal means in a fitted `drtmle` object, in addition to contrasts between
those means. By default the method gives confidence intervals about each mean.

@@

The `contrast` option also accepts as input a list of functions to be used to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald-style confidence intervals for ratios, it is common to
compute the interval on the logarithmic scale and subsequently back-transform.
That is, the confidence interval takes the form
\begin{equation*}
  \exp \biggl[\log \biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
\end{equation*}
where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
the delta method, an estimate of this standard error is given by
\begin{equation*}
  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
\end{equation*}
where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
`drtmle` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$. This
confidence interval may be obtained as follows.

@@

More generally, this method of inputting the `contrast` option to the `ci`
method can be used to compute confidence intervals of the form
\begin{equation*}
  f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
  \nabla f(h(\psi_n)) \bigr\} \ ,
\end{equation*}
where $f$ (`contrast$f`) is the transformation of the confidence interval,
$f^{-1}$ (`contrast$f_inv`) is the back-transformation of the interval, $h$
(`contrast$h`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
(`contrast$fh_grad`) defines the gradient of the transformed contrast at the
estimated marginal means.

@@

Doubly-robust Wald-style hypothesis tests may be implemented using the
`wald_test` method. As with confidence intervals, there are three options for
testing. First, we may test the null hypothesis that the marginal means equal a
fixed value (or values). Below, we test the null hypothesis that $\psi_0(1) =
0.5$ and that $\psi_0(0) = 0.6$ by inputting a vector to the `null` option.

@@

As with the `ci` method, the `wald_test` method allows users to test linear
combinations of marginal means by inputting a vector of ones and negative ones
in the `contrast` option. We can use this to test the null hypothesis of no
average treatment effect or to test the null hypothesis that the average
treatment effect equals a given reference value, e.g., $0.05$, by specifying
the `null` option.

>>>>

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
    orcdi: 0000-0002-1019-8343
  - name: Nima S. Hejazi
    affiliation: University of California, Berkeley
    address: |
      | 2121 Berkeley Way,
      | Berkeley, California, 94704, U.S.A.
    email: \email{nhejazi@berkeley.edu}
    orcdi: 0000-0002-7127-2789
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \pkg{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect. The package additionally implements a targeted inverse probability of
  treatment weighted estimator that yields valid inference even when the
  probability of treatment is estimated via adaptive methods.
header-includes:
  - \usepackage{mathtools}
output: rticles::jss_article
bibliography: refs
---

@@

The two proposed procedures that yield doubly-robust inference are quite
involved, notably requiring iterative estimation of additional, complex nuisance
parameters. The \pkg{drtmle} package was motivated by the need for a
user-friendly tool to implement these methods. The package implements the TML
estimators of population-level effects proposed in \citet{vdl:2014:ijb} and
\citet{benkeser:2017:biometrika}, as well as several extensions including
cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
\pkg{drtmle} implements an IPTW estimator that overcomes shortcomings of
existing IPTW implementations with respect to statistical inference. In
particular, the estimator allows an adaptive estimator of the PS to be used in
construction of the IPTW estimator, while yielding valid statistical inference
about population-level effects. Both the TML and IPTW estimators implemented in
the \pkg{drtmle} package are capable of estimating population-level effects for
discrete-valued treatments; moreover, the package provides convenient utilities
for constructing confidence intervals and hypothesis tests about contrasts of
the mean outcome under different levels of treatment. This document explains
some of the key concepts underlying doubly-robust inference and illustrates
usage of the \pkg{drtmle} package.

@@

The main function of the package is the eponymous `drtmle()` function. This
function estimates the treatment-specific marginal mean for user-specified
levels of a discrete-valued treatment and computes a doubly-robust covariance
matrix for these estimates. The function implements either the
estimators of \citet{vdl:2014:ijb} (if `reduction = "bivariate"`) or the
improved estimators of \citet{benkeser:2017:biometrika}
(`reduction = "univariate"`). The package includes utility functions for
combining these estimates into estimates of average treatment effects, as well
as other contrasts, as discussed below.

The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
regression parameters may be estimated. Once the user has specified the data
arguments and how to estimate each regression parameter, the function proceeds
as follows:
  1. estimate the OR via the internal `estimateQ()` function;
  2. estimate the PS via the internal `estimateG()` function;
  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
     `estimategrn()` functions;
  4. iteratively modify initial estimates of the OR and PS via the
     `fluctuateQ()` and `fluctuateG()` functions until equations (5)-(7) are
     approximately solved;
  5. compute the TML estimates and covariance estimates based on the final OR.
We refer interested readers to \citet{benkeser:2017:biometrika} for a
theoretical derivation of how the iterative modification of initial OR and PS is
performed.

In addition to returning doubly-robust parameter and covariance estimates, the
`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly-robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we look at various
options for making calls to `drtmle()`.

## Estimating regressions
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle()` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated, then the estimators
computed by `drtmle()` achieve the nonparametric efficiency bound, resulting in
narrower confidence intervals and more powerful hypothesis tests. In order for
the estimates of the OR and PS to have the greatest chance of consistency, the
`drtmle()` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning
[@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
technique. We demonstrate each of these approaches in turn on a simple simulated
data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
2)` and $\psi_0(1)=$ `r round(EY1, 2)`.

@@

The code below illustrates how generalized linear models can be used to
estimate the regression parameters. In this call to `drtmle()`, we specify
`family = binomial()` so that logistic regression is used to estimate the OR.
We illustrate non-stratified regression for both the univariate (default) and
bivariate reductions below.

@@

Super learning is implemented in the R package \pkg{SuperLearner}
[@SuperLearnerPackage] and `drtmle()` provides the option to utilize super
learning for estimation of the OR, PS, and reduced-dimension regressions via the
`SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle()` either the
generalized linear model options for regression modeling or the super learner
options should be invoked. If both are called, the function will automatically
opt to using the super learner option. The inputs to the super learner options
are passed to the `SL.library` argument of the `SuperLearner()` function of the
\pkg{SuperLearner} package. We refer readers to the documentation of the
`SuperLearner()` function for information on how to properly specify a super
learner library. In the code below, we illustrate a call to `drtmle()` with a
relatively simple library. For the OR and PS, the library includes a main-terms
generalized linear model (via the function `SL.glm` from \pkg{SuperLearner}) and
an unadjusted generalized linear model (`SL.mean`). For the reduced-dimension
regressions we use a super learner library with a main-terms generalized linear
model (`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
\pkg{earth} package [@earthPackage]).

@@

The \pkg{drtmle} package also allows users to pass in their own functions for
estimating regression parameters via the `SL_` options. These functions must be
formatted properly for compatibility with `drtmle()`. The necessary formatting
is identical to the formatting required for writing a `SuperLearner` wrapper
(see `SuperLearner::write.SL.template()`). The \pkg{drtmle} package includes
`SL.npreg`, which is an example of how a user can specify a function for
estimating regression parameters. This functions fits a kernel regression
estimator using the \pkg{np} package [@npPackage]. Below we show the source code
for this function.

@@

In the above code, we see that a user-specified function should have options
`Y, X, newX, family, obsWeights, ...`, along with other options specific to the
function. The option `Y` corresponds to the outcome of the regression and `X`
to the variables onto which the outcome is regressed. The option `newX` should
be of the same class and dimension as `X` and is meant to contain new values of
variables at which to evaluate the regression fit. The `family` and
`obsWeights` options are not strictly necessary (indeed, they are not used by
`SL.npreg`), though these may be useful to include if one wishes to utilize the
function as part of a super learner library. The `SL.npreg` function proceeds by
checking whether there is sufficient variation in `Y` to justify fitting the
kernel regression (as specified by option `rangeThresh`), and if so, fits a
kernel regression with cross-validated bandwidth selection via the `npregbw()`
function. The output is then formatted in a specific way so that `drtmle()` is
able to retrieve the appropriate objects from the fitted regression. In
particular the output should be a list with named objects `pred` and `fit`; the
former including predictions made from the fitted regression on `newX`, the
latter including any information that may be needed to predict new values based
on the fitted object. A class is assigned to the `fit` object, so that an S3
`predict()` method may later be used to predict new values based on the fitted
regression object. The user must also specify this `predict()` method. Use
`drtmle:::predict.SL.npreg()` to view the simple `predict()` method for
`SL.npreg`.

In the code below, we demonstrate a call to `drtmle()` with `SL.npreg` used to
estimate each nuisance parameter. Here, we also illustrate the use of `stratify
= TRUE`, which fits a separate kernel regression of $Y$ onto $W$ in
observations with $A = 1$ and $A = 0$.

@@

Not surprisingly, averaging over multiple super learners can add considerable
computational expense to calls to `drtmle()`. The package is currently
parallelized at the level of individual calls to `SuperLearner()`. With
parallelization over enough cores, repeated super learning should not add too
much to the computational burden relative to a single call to `SuperLearner()`.

@@

The \pkg{drtmle} package also allows users to pass in their own estimated OR and
PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
user, `drtmle()` will ignore the nuisance parameter estimation routines
specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
noted in the package documentation, there is a specific format necessary for
inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
the OR evaluated at `A` and the observed values of `W`, with order determined by
the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
example below illustrates this concept.

@@

This section contains some more advanced mathematical material and may be
skipped over by users interested only in implementation. For simplicity, we
focus this discussion on the standard TML and AIPTW estimators. The ideas
generalize immediately to the DR-TML estimator, but the formulas are more
complex in this latter case.

@@

For TML estimators, it is standard practice to construct $D_i$ using the
*targeted* version of the OR (denoted by $\bar{Q}_n^{\star}$ above). Recall that
such estimates are obtained by modifying an initial OR estimate so as to ensure
a particular (set of) equation(s) is (are) solved.  However, simulations have
shown that better standard error estimates for TMLE may be achieved by instead
using the *initial* OR estimate in the computation of standard errors, as
described in the preceding subsection. To this end, The `drtmle()` function
includes the `targeted_se` option.

@@

Suppose we use $V$-fold cross-validation to obtain $V$ training-sample-specific
estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the
estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the
validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of
$Y_i$. We define $g_{n,i}^{cv}$ similarly. Using these estimates, we can define
\begin{equation*}
 D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} +
 \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1)
\end{equation*}
and compute an estimate of $\sigma^2_0$ using
\begin{equation*}
  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} -
  \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ .
\end{equation*}

In \pkg{drtmle}, cross-validated standard error estimates can be obtained by
setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1. For
example,

@@

Obtaining fully cross-validated standard errors can be fairly time-intensive
when super learner is used to estimate the OR and PS. In \pkg{drtmle}, we give
an option to compute standard error estimates that are *partially
cross-validated* and do not require any additional fitting beyond a standard
super learner. Here, we describe briefly how these estimates are obtained and
show calls to `drtmle()` for obtaining such estimates.

@@

Below, we demonstrate how to obtain such estimates using `drtmle()`.

@@

The \pkg{drtmle} package contains methods that compute doubly-robust confidence
intervals. The `ci()` method computes confidence intervals about the estimated
marginal means in a fitted object of S3 class `"drtmle"`, in addition to
contrasts between those means. By default, the method gives confidence intervals
about each mean.

@@

The `contrast` option also accepts as input a list of functions to be used to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald-style confidence intervals for ratios, it is common to
compute the interval on the logarithmic scale and subsequently back-transform.
That is, the confidence interval takes the form
\begin{equation*}
  \exp \biggl[\log \biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
\end{equation*}
where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
the delta method, an estimate of this standard error is given by
\begin{equation*}
  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
\end{equation*}
where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
This confidence interval may be obtained as follows.

@@

More generally, this manner of inputting the `contrast` option to the `ci()`
method can be used to compute confidence intervals of the form
\begin{equation*}
  f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
  \nabla f(h(\psi_n)) \bigr\} \ ,
\end{equation*}
where $f$ (`contrast$f`) is the transformation of the confidence interval,
$f^{-1}$ (`contrast$f_inv`) is the back-transformation of the interval, $h$
(`contrast$h`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
(`contrast$fh_grad`) defines the gradient of the transformed contrast at the
estimated marginal means.

@@

Doubly-robust Wald-style hypothesis tests may be implemented using the
`wald_test()` method. As with confidence intervals, there are three options for
testing. First, we may test the null hypothesis that the marginal means equal a
fixed value (or values). Below, we test the null hypothesis that $\psi_0(1) =
0.5$ and that $\psi_0(0) = 0.6$ by inputting a vector to the `null` option.

@@

As with the `ci()` method, the `wald_test()` method allows users to test linear
combinations of marginal means by inputting a vector of ones and negative ones
in the `contrast` option. We can use this to test the null hypothesis of no
average treatment effect or to test the null hypothesis that the average
treatment effect equals a given reference value, e.g., $0.05$, by specifying
the `null` option.
